Randomise papers

### 2023-11-27 Randomisation/Randomise_Papers.qmd
randomised second sample

---
title: Randomise papers
subtitle: Describes random selection of papers for inclusion in the review
author: Saoirse Kelleher
date: today
format: 
  html:
    code-fold: true
    theme: minty
bibliography: references.bib
---

@@

-   After reviewing this initial sample, the need for a **Second Sample** was identified as many of the papers included in the first did not represent our model of interest. This second, more narrowly targeted sample was drawn from papers citing @mackenzie2003, the original description of the model class which all subsequent implementations would be expected to cite. The same 4 year-window strata were used here, with 10 papers randomly drawn from each strata.

@@

# Repeat with the study metadata
studyDataSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                    sheet = "Study data", skip = 1) %>%
  # Cut out non-data rows
  slice(4:nrow(.)) %>%
  # Only need year data and ID reference
  select(Review_ID, Year)

@@

# Identify which papers qualify to be read
highlightedPapers <- rankedPapers %>%
  group_by(Query, YearPeriod) %>%
  mutate(PercentRank =  GroupRank/max(GroupRank)) %>%
  mutate(ToRead = case_when(
    # If there are fewer than 5, read them all
    max(GroupRank) <= 5 ~ "YES",
    # If there are 20 or fewer, read the top 5
    max(GroupRank) <= 20 & GroupRank <= 5 ~ "YES",
    max(GroupRank) <= 20 & GroupRank >= 5 ~ "NO",
    # If there are 21 or more, read the top 25%
    max(GroupRank) >= 21 & PercentRank <= 0.25  ~ "YES",
    max(GroupRank) >= 21 & PercentRank > 0.25 ~ "NO")) %>%
  ungroup() %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Review_ID, Group, OverallRank = randRank, GroupRank, ToRead)

@@

```{r re-add unranked papers}
unrankedPapers <- paperList %>% 
  # Drop empty rows, excluded, and incidental papers
  filter(is.na(Query) | Query == "INCIDENTAL" | Year == "EXCLUDE") %>%
  mutate(ToRead = "NO", Group = "Unranked", GroupRank = "NA", OverallRank = "NA") %>%
  select(Review_ID, Group, OverallRank, GroupRank, ToRead)

# Combine ranked and unranked papers
outputList <- rbind(highlightedPapers, unrankedPapers) %>%
  mutate(Review_ID = as.numeric(Review_ID)) %>%
  arrange(Review_ID)

# Write to disk
write_csv(outputList, file = "Randomisation/PrimarySampleRanks.csv")
```

@@

>>>>

---
title: Randomise papers
subtitle: Describes random selection of papers for inclusion in the review
author: Saoirse Kelleher
date: today
format: 
  html:
    code-fold: true
    theme: minty
    toc: true
bibliography: references.bib
---

@@

-   After reviewing this initial sample, the need for a **Second Sample** was identified as many of the papers included in the first did not represent our model of interest. This second, more narrowly targeted sample was drawn from papers citing @mackenzie2003, the original description of the model class which all subsequent implementations would be expected to cite. Papers were acquired via Web of Science in September 2023. The same 4 year-window strata were used here, with 15 papers randomly drawn from each strata. A fifth strata was added for 2022-2023, to capture the most recent papers on these models.

For transparency and reproducibility, the randomisation code is preserved in this document.

@@

# Repeat with the study metadata
studyDataSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                    sheet = "Study data", skip = 1) %>%
  # Cut out non-data rows
  slice(4:nrow(.)) %>%
  # Only need year data and ID reference
  select(Review_ID, Title, Year)

@@

# Identify which papers qualify to be read
highlightedPapers <- rankedPapers %>%
  group_by(Query, YearPeriod) %>%
  mutate(PercentRank =  GroupRank/max(GroupRank)) %>%
  mutate(ToRead = case_when(
    # If there are fewer than 5, read them all
    max(GroupRank) <= 5 ~ "YES",
    # If there are 20 or fewer, read the top 5
    max(GroupRank) <= 20 & GroupRank <= 5 ~ "YES",
    max(GroupRank) <= 20 & GroupRank >= 5 ~ "NO",
    # If there are 21 or more, read the top 25%
    max(GroupRank) >= 21 & PercentRank <= 0.25  ~ "YES",
    max(GroupRank) >= 21 & PercentRank > 0.25 ~ "NO")) %>%
  ungroup() %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Review_ID, Title, Group, OverallRank = randRank, GroupRank, ToRead)

@@

```{r re-add unranked papers}
unrankedPapers <- paperList %>% 
  # Drop empty rows, excluded, and incidental papers
  filter(is.na(Query) | Query == "INCIDENTAL" | Year == "EXCLUDE") %>%
  mutate(ToRead = "NO", Group = "Unranked", GroupRank = "NA", OverallRank = "NA") %>%
  select(Review_ID, Title, Group, OverallRank, GroupRank, ToRead)

# Combine ranked and unranked papers
outputList <- rbind(highlightedPapers, unrankedPapers) %>%
  mutate(Review_ID = as.numeric(Review_ID)) %>%
  arrange(Review_ID) %>%
  mutate(Sample = "First")

# Write to disk
write_csv(outputList, file = "Randomisation/FirstSampleRanks.csv")
```

@@

The list of papers pulled from Web of Science is loaded in - 1184 papers are represented here. Year-windows are assigned here as in the **First Sample.**

```{r Load second sample}
#| warning: false

# Load query and extract relevant columns
wosQuery <- readxl::read_xls("Queries/MacKenzie WOS/MacKenzie03.xls") %>%
  select(Review_ID, Title = `Article Title`, Year = `Publication Year`) %>%
  # Assign year period
  mutate(YearPeriod = case_when(Year %in% 2000:2005 ~ "2000:2005",
                                Year %in% 2006:2010 ~ "2006:2010",
                                Year %in% 2011:2015 ~ "2011:2015",
                                Year %in% 2016:2021 ~ "2016:2021",
                                Year %in% 2022:2023 ~ "2022:2023"))

head(wosQuery) %>%
  kable()
```

Papers are assigned a random number and ranked within their groups, and top-15 papers are flagged to be read. The output is saved for future reference.

```{r randomise second sample}

randomisedWOS <- wosQuery %>%
  # Add query flag to match with the first sample
  mutate(Query = "MacKenzie Cited") %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(GroupRank = row_number()) %>% 
  mutate(ToRead = case_when(
    GroupRank <= 15 ~ "YES",
    GroupRank >15 ~ "NO")) %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  mutate(Sample = "Second") %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Sample, Review_ID, Title, Group, OverallRank = randRank, GroupRank, ToRead)

write_csv(randomisedWOS, "Randomisation/SecondSampleRanks.csv")

head(randomisedWOS) %>%
  kable()
```

## Sample summary

A basic representation of the two sample's coverage is presented here.
